---
title: "media-misinfo"
author: "Isabelle Gibson"
format: html
editor: visual
execute: 
  message: false
  warning: false
  echo: false
---

```{r Load necessary libraries}

library(dplyr)
library(readr)
library(ggplot2)
library(broom)
library(janitor)
library(modelsummary)
library(kableExtra)

# Load the CSV file
df <- read_csv("anes_socialmedia_study.csv")
```

```{r Recode variables}
# Recode the dependent variable (trust in vote accuracy) to binary
df <- df %>%
  mutate(trust_vote = ifelse(w2trustvote >= 4, 1, 0))  # 1 = High trust, 0 = Low trust

# Filter out rows where w2fb1 is -7, -6, or -5, keeping perception_misinfo as the variable name
df <- df %>%
  filter(w2fb1 >= 1 & w2fb1 <= 5) %>%
  mutate(perception_misinfo = w2fb1)

# Filter out rows where w2fb4 is -7, -6, -5, -1, keep name pol_fbnews
df <- df %>%
  filter(w2fb4 >= 1 & w2fb4 <= 5) %>%
  mutate(pol_newsfb = w2fb4)
```

```{r Descriptive statistics for Key variables}

# Descriptive statistics for key variables
descriptive_stats <- df %>%
  summarise(
    N_trust_vote = n(),
    mean_trust_vote = mean(trust_vote, na.rm = TRUE),
    sd_trust_vote = sd(trust_vote, na.rm = TRUE),
    min_trust_vote = min(trust_vote, na.rm = TRUE),
    max_trust_vote = max(trust_vote, na.rm = TRUE),
    
    N_perception_misinfo = n(),
    mean_perception_misinfo = mean(perception_misinfo, na.rm = TRUE),
    sd_perception_misinfo = sd(perception_misinfo, na.rm = TRUE),
    min_perception_misinfo = min(perception_misinfo, na.rm = TRUE),
    max_perception_misinfo = max(perception_misinfo, na.rm = TRUE),
    
    N_pol_newsfb = n(),
    mean_pol_newsfb = mean(pol_newsfb, na.rm = TRUE),
    sd_pol_newsfb = sd(pol_newsfb, na.rm = TRUE),
    min_pol_newsfb = min(pol_newsfb, na.rm = TRUE),
    max_pol_newsfb = max(pol_newsfb, na.rm = TRUE)
  )

# Display the descriptive statistics with 15 column names
descriptive_stats %>%
  knitr::kable(col.names = c("Sample Size (Trust Vote)", "Mean (Trust Vote)", "SD (Trust Vote)", "Min (Trust Vote)", "Max (Trust Vote)",
                             "Sample Size (Perception of Misinformation)", "Mean (Perception of Misinformation)", "SD (Perception of Misinformation)", 
                             "Min (Perception of Misinformation)", "Max (Perception of Misinformation)", 
                             "Sample Size (Political News on FB)", "Mean (Political News on FB)", "SD (Political News on FB)", 
                             "Min (Political News on FB)", "Max (Political News on FB)"),
               caption = "Descriptive Statistics for Trust in Vote Accuracy, Perception of Misinformation, and Political News Exposure on Facebook") %>%
  kableExtra::kable_styling(full_width = F, bootstrap_options = c("striped", "hover"))
```

```{r Cross Tabs}
# Crosstab for Trust in Vote Accuracy by Perception of Misinformation
datasummary_crosstab(trust_vote ~ perception_misinfo, data = df, output = "dataframe")

#cross-tab for trust in vote accuracy and exposure to political news on FB
datasummary_crosstab(trust_vote ~ pol_newsfb, data = df, output = "dataframe")

```

```{r}
# Crosstab with renamed variables for better presentation: Trust_vote v. Perception_misinfo
datasummary_crosstab(
  `Trust in Vote Count Accuracy of the 2020 Election` ~ `Perception of Misinformation on Facebook`, 
  data = df %>%
    mutate(
      `Trust in Vote Count Accuracy of the 2020 Election` = ifelse(trust_vote == 1, "High Trust", "Low Trust"),
      `Perception of Misinformation on Facebook` = factor(perception_misinfo,
        levels = 1:5,
        labels = c("Not at all serious", "Slightly serious", "Moderately serious", "Very serious", "Extremely serious"))
    ),
  title = "Cross-Tab of Trust in Vote Count Accuracy (2020 Election) on Perception of Misinformation on Facebook",
  position = "H"
)

#cross tab for Trust in vote count by Exposure to Political News on FB 
datasummary_crosstab(
  `Trust in Vote Count Accuracy of the 2020 Election` ~ `Exposure to Political News on Facebook`, 
  data = df %>%
    mutate(
      `Trust in Vote Count Accuracy of the 2020 Election` = ifelse(trust_vote == 1, "High Trust", "Low Trust"),
      `Exposure to Political News on Facebook` = factor(pol_newsfb,
        levels = 1:5,
        labels = c("Never", "Some of the time", "About half the time", "Most of time", "Always"))
    ),
  title = "Cross-Tab for Trust in vote count by Exposure to Political News on Facebook",
  position = "H"
)
```

**Table 1: Trust in Vote Accuracy by Perception of Misinformation on Facebook**

This crosstab shows how trust in vote accuracy (low or high) varies with perception of misinformation on Facebook, ranging from "Not at all serious" (1) to "Extremely serious" (5).

Low Trust (0):

The majority of respondents with low trust in vote accuracy believe misinformation on Facebook is serious, with 31.1% perceiving it as "Very serious" (4) and 27.8% perceiving it as "Extremely serious" (5). This indicates that people with low trust in vote accuracy are more likely to view misinformation on Facebook as a significant problem.

Only 2.7% of respondents with low trust rated misinformation as "Not at all serious" (1).

High Trust (1):

Respondents with high trust in vote accuracy show a similar pattern but lean more heavily toward higher levels of concern about misinformation, with 35.3% perceiving it as "Very serious" (4) and 41.3% perceiving it as "Extremely serious" (5).

A very small percentage (1.1%) of respondents with high trust believe misinformation is "Not at all serious" (1).

Overall:

Across all respondents, 33.1% perceive misinformation as "Very serious" (4) and 34.3% as "Extremely serious" (5), indicating that a large proportion of the total sample believes misinformation is a serious problem. Trust in vote accuracy (both low and high) is concentrated at these higher levels of perceived misinformation.

Table 2: Trust in Vote Accuracy by Exposure to Political News on Facebook

This cross-tab shows how trust in vote accuracy (low or high) varies with exposure to news on Facebook, ranging from "Never" to "Always."

Low Trust (0):

Among respondents with low trust, the largest group (33%) reports exposure to news on Facebook "Some of the time," followed by "About half of the time" (22.4%) and "Most of the time" (22.4%).

A smaller percentage (11.6%) reports never seeing political news on Facebook.

High Trust (1):

Similarly, respondents with high trust in vote accuracy also tend to fall in the "Some of the time" category (32.8%). However, a slightly larger proportion (24%) reports seeing news "Most of the time" compared to those with low trust.

The percentage of high-trust respondents who never see news on Facebook is similar to that of low-trust respondents (12.2%).

Overall:

Across all respondents, 32.9% report seeing news on Facebook "Some of the time," with smaller percentages at "About half of the time" (21.1%) and "Most of the time" (23.2%).

The distribution of exposure to news is relatively even across trust levels, with no extreme differences between those who report low or high trust.

Summary of Findings:

Perception of Misinformation: Both low and high trust respondents tend to view misinformation as a serious issue, but those with high trust lean more heavily toward the highest levels of perceived seriousness.

Exposure to News on Facebook: There is no drastic difference in exposure between low-trust and high-trust respondents, although high-trust respondents are slightly more likely to report seeing news "Most of the time" compared to low-trust respondents.

These crosstabs show that perception of misinformation has a clearer relationship with trust in vote accuracy than exposure to news, as both high and low trust respondents similarly report moderate levels of news exposure.

```{r}
#Bar graph for Trust in Vote Accuracy by Exposure to Political News on Facebook

ggplot(df, aes(x = as.factor(pol_newsfb), fill = as.factor(trust_vote))) +
  geom_bar(position = "fill") +
  labs(title = "Trust in Vote Accuracy by Exposure to Political News on Facebook",
       x = "Exposure to Political News on Facebook",
       y = "Proportion of Trust in Vote Accuracy",
       fill = "Trust in Vote Accuracy") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = c("1" = "Never", "2" = "Rarely", "3" = "Sometimes", 
                              "4" = "Often", "5" = "Always")) +
  theme_minimal(base_size = 12)


# Bar graph for Trust in Vote Accuracy by Perception of Misinformation
ggplot(df, aes(x = as.factor(perception_misinfo), fill = as.factor(trust_vote))) +
  geom_bar(position = "fill") +
  labs(title = "Trust in Vote Accuracy by Perception of Misinformation on Facebook",
       x = "Perception of Misinformation on Facebook",
       y = "Proportion of Trust in Vote Accuracy",
       fill = "Trust in Vote Accuracy") +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_x_discrete(labels = c("1" = "Not at all serious", "2" = "Slightly serious", 
                              "3" = "Moderately serious", "4" = "Very serious", 
                              "5" = "Extremely serious")) +
  theme_minimal(base_size = 12)

```
